Cerebral oxygenation manuscript figures

### 2023-06-12 Scripts/R/markdown_files/archive/cm_follow_up_figures.Rmd
12 June 2023 clean and organize scripts for publishing

---
title: "Cerebral malaria follow-up visit data"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: html_document
---

@@

load("all_dfa_results.Rdata")

@@

  dplyr::filter(grepl("avg_Hb", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "THC", "O2_sat")) %>% 
  mutate(measurement = factor(measurement, levels = c("O2_sat", "THC"))) %>% 
  mutate(pvals = ifelse(measurement == "THC", round(thc_fu_pval, 3), round(o2sat_fu_pval, 3))) %>% 
  mutate(ycoord = ifelse(measurement == "THC", 400, 100)) %>%

@@

  dplyr::filter(!grepl("avg|second", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "THC", "O2_sat")) %>% 
  mutate(measurement = factor(measurement, levels = c("O2_sat", "THC"))) %>% 
  mutate(pvals = ifelse(measurement == "THC", round(thc_alpha_fu_pval, 3), round(o2sat_alpha_fu_pval, 3))) %>%

>>>>

---
title: "Cerebral malaria follow-up visit data"
author: "Rachel Smith"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

base_dir <- "/Users/work/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Ackerman Lab 2019-2021/cerebral_oxygenation/"

load(paste0(base_dir, "objects/all_dfa_results.Rdata"))

@@

  dplyr::filter(grepl("avg_Hb", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "Hbtot", "Hboxy")) %>% 
  mutate(measurement = factor(measurement, levels = c("Hboxy", "Hbtot"))) %>% 
  mutate(pvals = ifelse(measurement == "Hbtot", round(thc_fu_pval, 3), round(o2sat_fu_pval, 3))) %>% 
  mutate(ycoord = ifelse(measurement == "Hbtot", 400, 100)) %>%

@@

  dplyr::filter(!grepl("avg|second", variable)) %>% 
  mutate(measurement = ifelse(grepl("conc", variable), "Hbtot", "Hboxy")) %>% 
  mutate(measurement = factor(measurement, levels = c("Hboxy", "Hbtot"))) %>% 
  mutate(pvals = ifelse(measurement == "Hbtot", round(thc_alpha_fu_pval, 3), round(o2sat_alpha_fu_pval, 3))) %>%

### 2023-06-12 Scripts/R/markdown_files/archive/final_figures.Rmd
12 June 2023 clean and organize scripts for publishing

# Load libraries
library(tidyverse)
library(readxl)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)

# Load data
load("objects/glm_model.Rdata")
load("objects/bandpass_filtered_dfa_results.Rdata")
load("objects/sims.Rdata")

@@

df_hbox <- read_csv("Data/Hans_datatable_exports/malawi key data v04Jan2022.csv") %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

@@

# number of subjects 
df_status <- df_hbox %>% 
  count() %>% 
  pivot_wider(1:2, names_from = "status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  count(sex) %>% 
  dplyr::filter(sex != "") %>% 
  pivot_wider(names_from = "sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>%

  pivot_wider(c(1,4), names_from = "status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

@@

#### A. Correlations of variables within subject groups
```{r fig3a, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

# calculate spearman correlations in each group

f_calc_cor <- function(df, group) {

  m <- df %>%
    dplyr::filter(status == group) %>% 
    dplyr::select(-c(subject_id, status)) %>% 
    cor(method = "spearman")

  diag(m) <- 0

  m
}

# create run correlations of potential regressors
m_cor_um <- df_model_impute %>% f_calc_cor(group = "UM")

m_cor_cm <- df_model_impute %>% f_calc_cor(group = "CM")

# correlation matrix

# write plot function
  f_plot_cor <- function(mat) {

  melt(mat) %>%

    mutate_if(is.factor, factor, levels = c("hct", "lactate", "cerebral_hb_oxy", "cerebral_hb_tot", "cerebral_hb_oxy_alpha2", "cerebral_hb_tot_alpha2")) %>%

    ggplot(aes(x = X1, y = X2, fill = value, label = ifelse(value != 0, round(value, 2), NA))) + 
    geom_tile() +
    scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
    geom_text(size = 9) +

    xlab("") +
    ylab("") +
    theme_bw() +
    theme(# text = element_text(family = "Arial"),
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text = element_text(size = 35),
          legend.text = element_text(size = 40),
          legend.title = element_text(size = 50),
          legend.key.size = unit(2, 'cm'),
          plot.title = element_text(size = 30)
    )

  # extract legend
  p_leg <- m_cor_um %>% f_plot_cor() %>% get_legend()

  # create plots
  m_cor_um %>% f_plot_cor() + ggtitle("Uncomplicated malaria") + theme(legend.position = "none")

  m_cor_cm %>% f_plot_cor() + ggtitle("Cerebral malaria") + theme(legend.position = "none")

  # as_ggplot(p_leg)

@@

  # plot patients
ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status), size = 3) +
  stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, fill = status),
                    level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  # stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, shape = status),
  #           na.rm = FALSE) +

@@

df_brain_swell_models %>% write.csv(file = "outputs/brain_swell_models.csv")

@@

df_figs1a <- read.table("Data/signal_segments/Hb_tot/TM0001CM01.txt")

@@

df_figs1b <- read.table("Data/filtered_signals/0.2s_mean_0.001_1_filt/Hb_tot/TM0001CM01.txt") %>% mutate(time = row_number())

@@

  lims(y = c(-0.001, 0.001)) +

  labs(y = "Bandpass filtered signal", x = "Time") +
  ggtitle("Pre-processed signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        axis.text = element_blank(),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

@@

# create color vector
my_col <- c("white" = "#00BFC4", "brown" = "#7CAE00", "pink" = "#F8766D")

@@

  ggplot(aes(x = noise, y = alpha)) +
  geom_violin(aes(color = noise)) +
  geom_jitter(position = position_jitter(0.2), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = noise), size = 0.2, width = 0.5) +

  scale_color_manual(values = my_col) +
  labs(x = "noise type") +
  scale_y_continuous(limits = c(0, 1.7)) +
  theme_bw() +
  theme(legend.position = "none")  +

>>>>

# Load libraries
library(tidyverse)
library(readxl)
library(DescTools)
library(janitor)
library(FSA)
library(factoextra)
library(ggrepel)
library(ggpubr)
library(tidymodels)
library(vip)
library(scales)
library(reshape)
library(extrafont)
library(grDevices)
library(psych)

# Load data
base_dir <- "/Users/ray/Library/CloudStorage/OneDrive-NationalInstitutesofHealth/Ackerman Lab 2019-2021/cerebral_oxygenation/"

load(paste0(base_dir, "objects/20221125_final_glm_models.Rdata"))
load(paste0(base_dir, "objects/glm_model.Rdata"))
load(paste0(base_dir, "objects/brain_bandpass_filtered_dfa_results.Rdata"))
load(paste0(base_dir, "objects/sims.Rdata"))

# ALT LOADING: these objects are located in a directory 3 levels above, i.e.:
load("../../../objects/20221125_final_glm_models.Rdata")

@@

df_hbox <- read_csv(paste0(base_dir, "Data/Hans_datatable_exports/malawi key data v04Jan2022.csv")) %>% 
  as_tibble() %>% 
  janitor::clean_names() %>% 
  mutate(status = replace(status, status == "HV", "HC")) %>% 
  dplyr::rename("subject_id" = "subject_id_session") %>%

@@

# number of subjects 
df_status <- df_hbox %>% 
  count() %>% 
  pivot_wider(names_from = "status", values_from = "n") %>% 
  mutate(variable = "Number of subjects") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

# Sex
df_sex <- df_hbox %>% 
  count(sex) %>% 
  pivot_wider(id_cols = "status", names_from = "sex", values_from = "n") %>% 
  mutate(perc_fem = round(Female/(Female + Male)*100, 2)) %>% 
  dplyr::select(-c(Female, Male)) %>%

  pivot_wider(names_from = "status", values_from = "perc_fem") %>% 
  mutate(variable = "Sex (% female)") %>% 
  dplyr::select(variable, everything()) %>% 
  mutate_if(is.numeric, as.character)

@@

#### A. Correlations of variables within subject groups
```{r fig3a_cors}

# find pearson correlation across all variables of interest, including p-values and adjusted p-values

cm_cor <- df_model %>% 
  dplyr::filter(status == "CM") %>% 
  dplyr::select(-c(subject_id, status)) %>% 
  corr.test(method = "pearson", adjust = "none")

um_cor <- df_model %>% 
  dplyr::filter(status == "UM") %>% 
  dplyr::select(-c(subject_id, status)) %>% 
  corr.test(method = "pearson", adjust = "none")

# combine into tibble
df_fig3a <- cm_cor$r %>% 
  as_tibble %>% 
  mutate(var1 = colnames(.), .before = hct) %>% 
  pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>% 
  mutate(status = "CM", .before = var1) %>%

  left_join(

    cm_cor$p %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "p_value"),

    by = c("var1", "var2")

  ) %>%

  # adjust p-values
  mutate(adj_p_val = p.adjust(p_value, method = "BH")) %>%

    um_cor$r %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>% 
      mutate(status = "UM", .before = var1) %>%

      left_join(

        um_cor$p %>% 
          as_tibble %>% 
          mutate(var1 = colnames(.), .before = hct) %>% 
          pivot_longer(2:ncol(.), names_to = "var2", values_to = "p_value"),

        by = c("var1", "var2")

      ) %>% 
      mutate(adj_p_val = p.adjust(p_value, method = "BH"))

  ) %>%

  # set factor levels for variables
  mutate_at(2:3, factor, levels = c("hct", "lactate", "cerebral_hb_oxy", "cerebral_hb_tot", "cerebral_hb_oxy_alpha2", "cerebral_hb_tot_alpha2")) %>%

  # set diagonal equal to zero
  mutate(pearsons_r = ifelse(var1 == var2, 0, pearsons_r))

```{r fig3a_sig_cors}

### find lowest |r| value that corresponds to a p < 0.05

t_stat_cor <- function(r, n){(r*sqrt(n - 2)) / (sqrt(1 - r^2))} # equation to calculate t-stat given correlation and n

## for CM:

# combine pearson's r and sample size
df_t_cor_cm <- cm_cor$r %>% 
  as_tibble %>% 
  mutate(var1 = colnames(.), .before = hct) %>% 
  pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>%

  left_join(

    cm_cor$n %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "n")

  ) %>%

  # calculate t-stat of each
  mutate(t = ifelse(pearsons_r != 1, t_stat_cor(pearsons_r, n), NA))

# compare with t-distribution table
lowest_sig_r_cm <- read_excel(paste0(base_dir, "Data/t_distribution_table.xlsx")) %>% 
  dplyr::rename("n" = "One Tail") %>% 
  filter(!(n %in% c("Two Tails", "df"))) %>% 
  dplyr::select(1:2) %>% 
  mutate(n = as.numeric(n)) %>%

  left_join(df_t_cor_cm, by = "n") %>% 
  filter(!is.na(t)) %>%

  # find significance by t-value
  mutate(is_sig = ifelse(abs(t) > `0.05`, "yes", "no")) %>%

  # pull only significant correlations
  filter(is_sig == "yes") %>%

  # find lowest r value associated with significant correlation
  slice(which.min(abs(pearsons_r))) %>% pull(pearsons_r)

## repeat for UM:

# combine pearson's r and sample size
df_t_cor_um <- um_cor$r %>% 
  as_tibble %>% 
  mutate(var1 = colnames(.), .before = hct) %>% 
  pivot_longer(2:ncol(.), names_to = "var2", values_to = "pearsons_r") %>%

  left_join(

    um_cor$n %>% 
      as_tibble %>% 
      mutate(var1 = colnames(.), .before = hct) %>% 
      pivot_longer(2:ncol(.), names_to = "var2", values_to = "n")

  ) %>%

  # calculate t-stat of each
  mutate(t = ifelse(pearsons_r != 1, t_stat_cor(pearsons_r, n), NA))

# compare with t-distribution table
lowest_sig_r_um <-  read_excel(paste0(base_dir, "Data/t_distribution_table.xlsx")) %>% 
  dplyr::rename("n" = "One Tail") %>% 
  filter(!(n %in% c("Two Tails", "df"))) %>% 
  dplyr::select(1:2) %>% 
  mutate(n = as.numeric(n)) %>%

  left_join(df_t_cor_um, by = "n") %>% 
  filter(!is.na(t)) %>%

  # find significance by t-value
  mutate(is_sig = ifelse(abs(t) > `0.05`, "yes", "no")) %>%

  # pull only significant correlations
  filter(is_sig == "yes") %>%

  # find lowest r value associated with significant correlation
  slice(which.min(abs(pearsons_r))) %>% pull(pearsons_r)

```{r fig3a}

df_fig3a %>%

  ggplot(aes(x = var1, y = var2, fill = pearsons_r, label = ifelse(pearsons_r != 0, round(pearsons_r, 2), NA))) + 
  geom_tile(aes(width = 0.9, height = 0.9)) +

  facet_wrap(~ status) +

  scale_fill_gradient2(low = "#4575b4", high = "#d73027") +
  geom_text(size = 4) +

  xlab("") +
  ylab("") +
  theme_classic() +
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 0.95))

```
In the CM group (n = 46), absolute correlation values greater 0.32 are statistically significant at a threshold of p < 0.05. In the UM group (n = 36), absolute correlation values greater than the absolute value of 0.43 are statistically significant.
\newpage

@@

  # plot patients
ggplot() +
  geom_point(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status), size = 3) +
  # stat_conf_ellipse(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, fill = status),
  #                   level = 0.95, npoint = 100, bary = TRUE, alpha = 0.1, geom = "polygon") +
  # stat_mean(data = df_pnts, mapping = aes(x = PC1, y = PC2, color = status, shape = status),
  #           na.rm = FALSE) +

@@

df_brain_swell_models %>% write.csv(file = paste0("outputs/brain_swell_models.csv"))

@@

df_figs1a <- read.table(paste0(base_dir, "Data/signal_segments/Hb_tot/TM0001CM01.txt"))

@@

df_figs1b <- read.table(paste0(base_dir, "Data/filtered_signals/0.2s_mean_0.001_1_filt/Hb_tot/TM0001CM01.txt")) %>% 
  mutate(time = row_number())

@@

  #lims(y = c(-0.001, 0.001)) +

  labs(y = "Bandpass filtered signal", x = "Time") +
  ggtitle("Pre-processed signal") +
  theme_bw() +
  theme(# text = element_text(family = "Arial"),
        #axis.text = element_blank(),
        axis.title = element_text(size = 25),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))

@@

# create color vector
my_col <- c("white" = "white", "brown" = "sienna2", "pink" = "pink")

@@

  ggplot(aes(x = noise, y = alpha, color = noise)) +
  geom_violin(fill = "transparent") +
  geom_jitter(position = position_jitter(width = 0.1), shape = 1, size = 2) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = noise), size = 0.2, width = 0.5) +

  scale_color_manual(values = my_col) +
  labs(x = "noise type", y = "alpha (\u03b1)") +
  scale_y_continuous(limits = c(0, 1.7)) +
  theme_dark() +
  theme(legend.position = "none")  +

### 2023-06-12 Scripts/R/markdown_files/archive/tidymodels_analysis.Rmd
12 June 2023 clean and organize scripts for publishing

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC1: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +

>>>>

  # design
  xlab(paste0("PC1: ", round(pca_eig_val$variance.percent[1], 2), "% of variance")) +
  ylab(paste0("PC2: ", round(pca_eig_val$variance.percent[2], 2), "% of variance")) +
